(Re-)bridging the anatomy ontologies with SSSOM

### 2024-03-31 sssomt.tex
Suggested corrections from N. Harris.

One of the pre-requisites for the exploitation of datasets from
high-throughput experimental methods, such as single-cell RNA sequencing
datasets, beyond their lab of origin is the ability to annotate samples
and results with consistent terms. That is, we need controlled
vocabularies to properly record information such as the exact
techniques used, the organs or tissues from which the experimental
samples were taken, or the cell types that were identified. Several
ontologies have been developed for this kind of purpose, such as the
Experimental Factor Ontology (EFO) to annotate experimental
methods~\cite{malone2010}, the Drosophila Anatomy Ontology (FBbt) to
annotate anatomical structures and cell types in
\species{D.~melanogaster}~\cite{costa2013}, the Worm Anatomy Ontology
(WBbt) to do the same in \species{C.~elegans}~\cite{lee2003}, etc. Most
of those ontologies are grouped under the umbrella of the OBO
Foundry~\cite{jackson2021}.

>>>>

One of the prerequisites for the reuse and renalysis of datasets from
high-throughput experimental methods --~such as single-cell RNA sequencing
datasets~-- beyond their lab of origin is the annotation of samples
and results with standardised terms. That is, we need controlled
vocabularies to properly record information such as the exact
techniques used, the organs or tissues from which the experimental
samples were taken, or the cell types that were identified. Several
ontologies have been developed for this kind of purpose, such as the
Experimental Factor Ontology (EFO) to annotate experimental
methods~\cite{malone2010}, the Drosophila Anatomy Ontology (FBbt) to
annotate anatomical structures and cell types in
\species{D.~melanogaster}~\cite{costa2013}, the Worm Anatomy Ontology
(WBbt) to do the same in \species{C.~elegans}~\cite{lee2003}, etc. Most
of those ontologies are grouped under the umbrella of the OBO
Foundry~\cite{jackson2021}.
